---
title: "FINAL_PROJECT"
author: "rama krishna"
date: "12/9/2021"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}
library(tidyverse)
library(cluster)
library(factoextra)
setwd("C:/Users/krish/OneDrive/Desktop/R_MLCODES/64060_final project")
finaldata <- read.csv("ML_Project-Data.csv")
head(finaldata)
```
# Modify the column names for easy reference 
```{r}
colnames(finaldata) <- c('major', 'starting_salary', 'midcareer_salary', 'career_growth_inpercentage' , 'percent10_salary' , 'percent25_salary', 'percent75_salary' , 'percent90_salary')
#View(finaldata)
```
#remove the dollar sign from the data .
```{r}
majors <- finaldata['major']
majors
salary <- finaldata %>%
          select(-major) %>%
          mutate_all(function(x) as.numeric(gsub("[\\$,]", "",x))) %>% 
          mutate(career_growth_inpercentage = career_growth_inpercentage/100)
a<- bind_cols(majors, salary)
head(a)
```
# determining the optimal number of clusters based on starting_salary , midcareer_salary, 10percent_salary, 90percent_salary.
```{r}
kdata <- a %>% 
  select(starting_salary, midcareer_salary,percent10_salary, percent90_salary) %>% scale()
head(kdata)
```
# determine the number of clusters by using elbow method
```{r}
fviz_nbclust(kdata,kmeans, method="wss") + labs(subtitle = "ELBOW METHOD")
```
# determine the number of clusters by using silhouette method 
```{r}
fviz_nbclust(kdata,kmeans, method="silhouette") + labs(subtitle = "silhouette method")
```
# determine the number of clusters of gap statistic method
```{r}
gap <- clusGap(kdata, FUN = kmeans, nstart = 25,
                    K.max = 10, B = 50)
fviz_gap_stat(gap)
```
# Set k equal to the optimal number of clusters which is 3 since k=3 from elbow method and gap statistics and run kmeans algorithm and visualize the clusters.
```{r}
set.seed(7)
num_clusters <- 3
k_means <- kmeans(kdata , num_clusters , iter.max = 15, nstart = 25)
a$clusters <- k_means[[1]]

salary_increment<- ggplot(a,
                        aes(x=starting_salary,
                            y=midcareer_salary,color=factor(clusters))) +
                  scale_x_continuous(labels=scales::dollar)+
                  scale_y_continuous(labels=scales::dollar)+
                  geom_point(alpha=4/5,size=6)+
                  labs(x="Starting Salary",y="Mid Career Salary",
                       title="Clustering by Starting Salary vs. Mid Career Salary",
                      colour="Clusters")+
                  scale_colour_manual(values=c("#EC2C73", "#29AEC7", "#FFDD30"))
# visualize the output
salary_increment
```
```{r}
ggplot(a, aes(starting_salary,midcareer_salary))+ geom_point(alpha = 0.6)+ 
  geom_smooth(se = F) +  #loses fit 
  scale_x_continuous(labels=scales::dollar)+
                  scale_y_continuous(labels=scales::dollar)
```
```{r}
paste('correlation coefficient',
      round(with(a, cor(starting_salary,midcareer_salary)), 4))
```
```{r}
salary_variation <- a %>% select(major,percent10_salary, percent25_salary, midcareer_salary, percent75_salary, percent90_salary, clusters) %>% gather(key = percentage, value = salary, -c(major, clusters))

salary_variation$percentage = factor(salary_variation$percentage,levels = c('percent10_salary', 'percent25_salary', 'midcareer_salary',               'percent75_salary', 'percent90_salary'))
```

```{r}
cluster_1 <- salary_variation %>% filter(clusters==1) %>%
          ggplot(aes(x=percentage,y=salary,group=major,color=major, order=salary))+
          geom_point()+
          geom_line()+
          labs(title="Cluster 1: Average Earners")+
          theme(axis.text.x = element_text(size=7,angle=25))
cluster_1
```
```{r}
cluster_2 <- salary_variation %>% filter(clusters==2) %>%
          ggplot(aes(x=percentage,y=salary,group=major,color=major, order=salary))+
          geom_point()+
          geom_line()+
          labs(title="Cluster 2: Slow Earners")+
          theme(axis.text.x = element_text(size=7,angle=25))
cluster_2
```

```{r}
cluster_3 <- salary_variation %>% filter(clusters==3) %>%
          ggplot(aes(x=percentage,y=salary,group=major,color=major, order=salary))+
          geom_point()+
          geom_line()+
          labs(title="Cluster 3: The Over Achievers")+
          theme(axis.text.x = element_text(size=7,angle=25))
cluster_3
```
# arranging the majors in the descending order of the career percentage growth.
```{r}
a <- a %>% arrange(desc(career_growth_inpercentage))
head(a,8)
```


