---
title: "assignment3"
author: "rama krishna"
date: "10/17/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
setwd("C:/Users/krish/OneDrive/Desktop/R_MLCODES/rmullapu_64060")

library(caret)
library(gmodels)
library(ISLR)
library(e1071)

Bank <- read.csv("UniversalBank.csv")

Bank$Personal.Loan <- as.factor(Bank$Personal.Loan)
Bank$CreditCard <- as.factor(Bank$CreditCard)
Bank$Online <- as.factor(Bank$Online)
```
# partitioning the data into training (60%) and validation (40%) sets.

```{r}
set.seed(100)

Bank_index <- createDataPartition(Bank$Age, p=0.4, list = F)
Bank_valid <- Bank[Bank_index,]
summary(Bank_valid)
Bank_train <- Bank[-Bank_index,]
summary(Bank_train)
```
{A}. Create a pivot table for the training data with Online as a column variable, CC as a row 
variable, and Loan as a secondary row variable. The values inside the table should convey 
the count. In R use functions melt() and cast(), or function table(). In Python, use panda 
dataframe methods melt() and pivot().

```{r}
# the pivot table should consists of only data of online, personal.loan, credit card and the ftable does not taking exclude = c(1:9)

Bank_valid1 <- Bank_valid[,c(10,13,14)]
summary(Bank_valid1)

Bank_train1 <- Bank_train[,c(10,13,14)]
summary(Bank_train1)

ftable(Bank_train1, row.vars = c(3,1), col.vars = "Online")
```
{B} Consider the task of classifying a customer who owns a bank credit card and is actively using 
online banking services. Looking at the pivot table, what is the probability that this customer 
will accept the loan offer? 

This is the probability of loan acceptance (Loan = 1) conditional on 
having a bank credit card (CC = 1) and being an active user of online banking services (Online 
= 1).

p(loan=1 | CC= 1, Online=1) = 57/ (469+57) = 0.10836501
```{r}
a = table(Bank_train1[,c(1,2,3)])
b<- as.data.frame(a)
b
View(b)
```
{C} Create two separate pivot tables for the training data. One will have Loan (rows) as a 
function of Online (columns) and the other will have Loan (rows) as a function of CC.

```{r}
ftable(Bank_train1, row.vars ="Personal.Loan", col.vars = "Online")
ftable(Bank_train1, row.vars ="Personal.Loan", col.vars = "CreditCard")
```
{D} Compute the following quantities [P(A | B) means “the probability ofA given B”]: 

i. A = P(CC = 1 | Loan = 1) (the proportion of credit card holders among the loan acceptors
    A= 0.30519

```{r}
m<-table(Bank_train[,c(14,10)])
m
View(m)
#S1<-m[2,2]/(m[2,2]+m[1,2])
#ANSWER:-> p(CC =1 | Loan = 1) = 94/(94+214) = 0.30519.
#S1
```
ii. B = P(Online = 1 | Loan = 1) = 0.607142
```{r}
n<- table(Bank_train[,c(13,10)])
View(n)
#ANSWER: -> p(Online =1 | Loan=1) = 187/(187+121) = 0.607142
```
iii. C = P(Loan = 1)   (the proportion of loan acceptors)

     C= p(Loan = 1) = (84+130+37+57)/(1904+214+787+94) = 308/2999
                   C= 0.1027009
                  
iv.D = P(CC = 1 | Loan = 0)  = 318/(318+757)
                           D = 0.295813
                        
v. E = P(Online = 1 | Loan = 0) = 1616/(1616+1075) = 0.6005203

vi. F =  P(Loan = 0) = (757+1147+318+469) /(1904+214+787+94) = 2691/ 2999
               F = 0.8972991
 
{E} Use the quantities computed above to compute the naive Bayes probability P(Loan = 1 | CC 
= 1, Online = 1).
```{r}
A=0.3052
B=0.6071
C=0.1027
D=0.2958
E=0.6005
F=0.8973
 NBP <- (A*B*C)/ ((A*B*C)+ (D*E*F))
 NBP 
```
{F} Compare this value with the one obtained from the pivot table in (B). Which is a more accurate estimate?

ANS:- The value obtained from pivot table in (B) is 0.10836501 and the value obtained from naive bayes probability is 0.106656.There is a very little difference between the values which is 0.00170901.The value of the pivot table is more accurate because pivot table does not assume that probabilities are independent.

{G} Which of the entries in this table are needed for computing P(Loan = 1 | CC = 1, Online = 1)? Run naive Bayes on the data. Examine the model output on training data, and find the entry that corresponds to P(Loan = 1 | CC = 1, Online = 1). Compare this to the number you obtained in (E)

```{r}
set.seed(150)
naivebayes <- naiveBayes(Personal.Loan~.,data=Bank_train1)
naivebayes
```
the output value that we got in the naivebayes caluculation is same as the value that in E..



